Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.Rmd
update the two plots with noisy index and some writing

The index function, up until this point, are all smooth, while this is not the case for all the index functions. `norm_kol`, a 1D projection function based on the Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{Y}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF). Denote the ECDF function as $F_{.}(u)$ with the subscript indicating either the projection or the random normal variable,  the `norm_kol` index is defined by

$$\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]$$

Figure \ref{fig:noisy-better-geo} compares the tracing plot of two optimisers: `search_geodesic` and `search_better`. This time, the interpolated path is no longer smooth when using either algorithm and  `search_geodesic` fails to optimise this index with barely any improvement of the index value. On the other hand, `search_better` is doing relatively well. With the theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] producing an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(nrow(boa5)), raw_data = boa5) %>% tail(1) %>% pull(index_val)`, `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 3)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`. A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 3)`]. At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumptions in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.

>>>>

The index function, up until this point, are all smooth, while this is not the case for all. A 1D projection function based on the Kolmogorov test, `norm_kol`, compares the projected data, $\mathbf{Y}_{n \times 1}$ to a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF). Let $F_{.}(u)$ be the ECDF function with the subscript $P$ and $u$ indicating the projected and randomly generated ECDF respectively,  the index is defined as:

$$\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]$$
With a more complex index function, it is interesting to understand 1) if all the three optimisers can reach the optimum, 2) when a local optimum is presented, whether the simulated annealing algorithm can escape that local optimum and find the global one.

```{r noisy-better-geo, fig.env = "widefigure", fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index. The trace plot shows the failure of pseudo derivative on non-smooth index. The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index."}
```

Figure \ref{fig:noisy-better-geo} presents the tracing plots of two optimisers: pseudo derivative and simulated annealing and as expected, the interpolated path is no longer smooth in either case. There is barely any improvement on the index value in the pseudo derivative algorithm while the simulated annealing algorithm has managed to get close to the index value of the theoretical best, indicated by the horizontal dashed line. In the PCA plot, simulated annealing is able to move gradually towards the theoretical best while the pseudo derivative method breaks after the initial few attempts.

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.tex
update the two plots with noisy index and some writing

The index function, up until this point, are all smooth, while this is
not the case for all the index functions. \texttt{norm\_kol}, a 1D
projection function based on the Kolmogorov test, compares the
difference between the 1D projected data, \(\mathbf{Y}_{n \times 1}\)
and a randomly generated normal distribution, \(y_n\) based on the
empirical cumulated distribution function (ECDF). Denote the ECDF
function as \(F_{.}(u)\) with the subscript indicating either the
projection or the random normal variable, the \texttt{norm\_kol} index
is defined by

\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\]

Figure \ref{fig:noisy-better-geo} compares the tracing plot of two
optimisers: \texttt{search\_geodesic} and \texttt{search\_better}. This
time, the interpolated path is no longer smooth when using either
algorithm and \texttt{search\_geodesic} fails to optimise this index
with barely any improvement of the index value. On the other hand,
\texttt{search\_better} is doing relatively well. With the theoretical
best basis {[}0, 1, 0, 0, 0{]} producing an index value of 0.17,
\texttt{search\_better} finds the final basis {[} 0.0376, -0.9916,
-0.0581, -0.0831, 0.0716{]} with an index value of 0.165. A further
polish step will give a marginal improvement of index value to 0.175
with a basis of {[} 0.0223, -0.9965, -0.0352, -0.0591, 0.0418{]}. At
this stage, the difference between the theoretical best and what has
been found is likely due to simulation error since the best possible
basis for a simulated data will be slightly off the theoretical best
basis, which is derived based on the distributional assumptions in
Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.

>>>>

The index function, up until this point, are all smooth, while this is
not the case for all. A 1D projection function based on the Kolmogorov
test, \texttt{norm\_kol}, compares the projected data,
\(\mathbf{Y}_{n \times 1}\) to a randomly generated normal distribution,
\(y_n\) based on the empirical cumulated distribution function (ECDF).
Let \(F_{.}(u)\) be the ECDF function with the subscript \(P\) and \(u\)
indicating the projected and randomly generated ECDF respectively, the
index is defined as:

\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\] With a more complex
index function, it is interesting to understand 1) if all the three
optimisers can reach the optimum, 2) when a local optimum is presented,
whether the simulated annealing algorithm can escape that local optimum
and find the global one.

\begin{Schunk}
\begin{widefigure}

{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}

\caption[1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index]{1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index. The trace plot shows the failure of pseudo derivative on non-smooth index. The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
\end{widefigure}
\end{Schunk}

Figure \ref{fig:noisy-better-geo} presents the tracing plots of two
optimisers: pseudo derivative and simulated annealing and as expected,
the interpolated path is no longer smooth in either case. There is
barely any improvement on the index value in the pseudo derivative
algorithm while the simulated annealing algorithm has managed to get
close to the index value of the theoretical best, indicated by the
horizontal dashed line. In the PCA plot, simulated annealing is able to
move gradually towards the theoretical best while the pseudo derivative
method breaks after the initial few attempts.
